---
title: "Gaussian Mixtures"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
pkgs = c("tidyverse","plotly", "shiny", "flexdashboard")
void = lapply(pkgs, library, character.only = TRUE)
```



Inputs {.sidebar}
==========================================================

```{r}
# sliderInput("eps", label = h3("Small epsilon to avoid singularities"), 
# min = 0, max = 0.2, value = 0.1)
```


```{r}
sliderInput("p", label = h3("Mixing probability p:"), 
min = 0, max = 1, value = 0.5)
```

```{r}
sliderInput("radius", label = h3("Radius: "), 
min = 0, max = 1, value = 0.5)

   sliderInput("angle", label = h3("Angle:"), 
min = 0, max = pi/2, value = pi/4)
```

```{r}
n_gridpoints = 100
eps = 0.1
tt <- expand_grid(angle = seq(eps, 1-eps, length.out = n_gridpoints),
                  radius = seq(eps, 1-eps, length.out = n_gridpoints),
                  p = c(0.5, seq(eps, 1-eps, length.out = n_gridpoints))) %>% 
  mutate(mu1 = sqrt((1-p)/p) * radius * sin(angle*pi/2),
         mu2 = -sqrt(p/(1-p)) * radius * sin(angle*pi/2),
         sigma1 = sqrt(1/p) * radius * cos(angle*pi/2),
         sigma2 = sqrt(1/(1-p) * (1-radius^2))) %>% 
  mutate(cumu3 = p * (mu1^3 + 3*mu1*sigma1^2) + (1-p) * (mu2^3 + 3*mu2*sigma2^2),
         cumu4 = p * (mu1^4 + 6*mu1^2*sigma1^2 + 3*sigma1^4) + (1-p) * (mu2^4 + 6*mu2^2*sigma2^2 + 3*sigma2^4) - 3)
```

Histograms
==================

Column {data-width=500}
-----------------------------------------------------------------------

### Histogram

<!--
> The red lines visualize the mean 0 and standard deviation of one of the Gaussian mixture.
> The blue and green solid lines denote the mean of the two summands.
-->

```{r}
renderPlot({
  dd = bmixture::rmixnorm(n = 10000, 
                          weight = c(input$p, 1 - input$p), 
                          mean = c(sqrt((1-input$p)/input$p) * input$radius * sin(input$angle), 
                                   -sqrt((1-input$p)/input$p) * input$radius * sin(input$angle)),
                          sd = c(sqrt(1/input$p) * input$radius * cos(input$angle),
                                 sqrt(1/(1-input$p) * (1-input$radius^2))))
  
  mu1 = sqrt((1-input$p)/input$p) * input$radius * sin(input$angle)
  mu2 = -sqrt(input$p/(1-input$p)) * input$radius * sin(input$angle)
  sigma1 = sqrt(1/input$p) * input$radius * cos(input$angle)
  sigma2 = sqrt(1/(1-input$p) * (1-input$radius^2))

  
  ggplot(NULL, aes(x = dd)) + geom_histogram() + 
        geom_vline(xintercept = 0, size = 3, color = "red") + geom_vline(xintercept = 1, size = 2, color = "red", linetype = 2) + geom_vline(xintercept = -1, size = 2, color = "red", linetype = 2) +
    geom_vline(xintercept = mu1, size = 2, color = "blue") + geom_vline(xintercept = mu1 + sigma1, size = 1, color = "blue", linetype = 2) + geom_vline(xintercept = mu1 - sigma1, size = 1, color = "blue", linetype = 2) + 
    geom_vline(xintercept = mu2, size = 2, color = "green") + geom_vline(xintercept = mu2 + sigma2, size = 1, color = "green", linetype = 2) + geom_vline(xintercept = mu2 - sigma2, size = 1, color = "green", linetype = 2) 
    
  
  # hist(, 
  #      freq = FALSE, 
  #      breaks = 30, 
  #      col = "gray" )
})  
```

### Minima and Maxima of Third Cumulant (fixed p, across all angles and radii)

```{r}
plotly::renderPlotly({
  plotly::ggplotly(
    ggplot(tt %>% 
             filter(p >= eps & p <= 1-eps) %>%
             group_by(p) %>%
             nest() %>%
             mutate(cumu3_min = map_dbl(data, ~min(.x$cumu3))) %>%
             mutate(cumu3_max = map_dbl(data, ~max(.x$cumu3))) %>%
             mutate(cumu4_min = map_dbl(data, ~min(.x$cumu4))) %>%
             mutate(cumu4_max = map_dbl(data, ~max(.x$cumu4))) %>%
             ungroup() %>%
             select(-data) %>%
             pivot_longer(starts_with("cumu")),
           aes(x = p, y = value, col = name)) + geom_line())})
```




Column {data-width=400}
-----------------------------------------------------------------------

### Details on Parametrisation


The density of a Gaussian mixture is of the form
$$
p \cdot \frac{1}{\sqrt{2 \pi \sigma_1^2}} e^{-\frac{1}{2} \left( \frac{x-\mu_1}{\sigma_1} \right)^2} + 
(1-p) \cdot \frac{1}{\sqrt{2 \pi \sigma_2^2}} e^{-\frac{1}{2} \left( \frac{x-\mu_2}{\sigma_2} \right)^2}
$$

Since $\left( r, \theta \right)$ represents $\sqrt{p} \sigma_2$ and $\sqrt{\frac{p}{1-p}} \mu_1$ in polar coordinates, we have that 
$$
\left(\mu_1, \sigma_1\right) = \left( \sqrt{\frac{1-p}{p}} \cdot r \cdot \sin(\theta) , \sqrt{\frac{1}{p}} \cdot r \cdot \cos(\theta)\right).
$$
Note that $\left(\mu_2, \sigma_2 \right)$ are a function of $\left(\mu_1, \sigma_1 \right)$ to ensure that $\mathbb{E}(X) = 0$ and $\mathbb{V}(X) = 1$. 
We have 
$$
\left(\mu_2, \sigma_2 \right) = \left( -\frac{p}{1-p}\mu_1, \sqrt{ \frac{1}{1-p} \left[1 -\left(p \sigma_1^2 + \frac{p}{1-p}\mu_1^2\right)\right]} \right).
$$

Substituting for all relevant quantities in terms of $\left( p, r, \theta \right)$, we obtain that 
$$
\left(\theta, r, p \right) \mapsto 
\begin{pmatrix}
p \\
\mu_1 \\
\mu_2 \\
\sigma_1 \\
\sigma_2
\end{pmatrix}
=
\begin{pmatrix}
p \\
\sqrt{\frac{1-p}{p}} \cdot r \cdot \sin(\theta) \\
-\sqrt{\frac{1-p}{p}} \cdot r \cdot \sin(\theta) \\
\sqrt{\frac{1}{p}} \cdot r \cdot \cos(\theta) \\
\sqrt{\frac{1}{1-p}} \cdot \sqrt{1-r^2}
\end{pmatrix}
$$
<!-- $$ -->
<!-- \left(\theta, r, p \right) \mapsto  -->
<!-- \begin{pmatrix} -->
<!-- p \\ -->
<!-- \mu_1 \\ -->
<!-- \mu_2 \\ -->
<!-- \sigma_1 \\ -->
<!-- \sigma_2 -->
<!-- \end{pmatrix} -->
<!-- = -->
<!-- \begin{pmatrix} -->
<!-- p \\ -->
<!-- \sqrt{\frac{1-p}{p}} \cdot r \cdot \sin(\theta) \\ -->
<!-- -\sqrt{\frac{1-p}{p}} \cdot r \cdot \sin(\theta) \\ -->
<!-- \sqrt{\frac{1}{p}} \cdot r \cdot \cos(\theta) \\ -->
<!-- \sqrt{\frac{1}{1-p}} \cdot \sqrt{1-r^2} -->
<!-- \end{pmatrix} -->
<!-- $$ -->

The cumulants are calculated as 

$$
\kappa^{(3)} = p \cdot (\mu_1^3 + 3 \cdot \mu_1 \cdot \sigma_1^2) + (1-p) \cdot (\mu_2^3 + 3 \cdot \mu_2 \cdot \sigma_2^2)
$$
and

$$
\kappa^{(4)} = p  \cdot  (\mu_1^4 + 6 \cdot \mu_1^2 \cdot \sigma_1^2 + 3 \cdot \sigma_1^4) + (1-p)  \cdot  (\mu_2^4 + 6 \cdot \mu_2^2 \cdot \sigma_2^2 + 3 \cdot \sigma_2^4) - 3
$$

Column {data-width=100}
-----------------------------------------------------------------------

### Value of $p$

```{r}
renderValueBox({
  valueBox(value = round(input$p, 4))
})
```

### Value of $\mu_1$ and $\mu_2$

```{r}
renderValueBox({
  valueBox(value = round(sqrt((1-input$p)/input$p) * input$radius * sin(input$angle), 4))
})

renderValueBox({
  valueBox(value = round(-sqrt(input$p/(1-input$p)) * input$radius * sin(input$angle), 4))
})
```

### Value of $\sigma_1$ and $\sigma_2$

```{r}
renderValueBox({
  valueBox(value = round(sqrt(1/input$p) * input$radius * cos(input$angle), 4))
})

renderValueBox({
  valueBox(value = round(sqrt(1/(1-input$p) * (1-input$radius^2)), 4))
})
```

### Value of $\kappa^{(3)}$ and $\kappa^{(4)}$

```{r}
renderValueBox({
  mu1 = sqrt((1-input$p)/input$p) * input$radius * sin(input$angle)
  mu2 = -sqrt(input$p/(1-input$p)) * input$radius * sin(input$angle)
  sigma1 = sqrt(1/input$p) * input$radius * cos(input$angle)
  sigma2 = sqrt(1/(1-input$p) * (1-input$radius^2))

  valueBox(value = round(input$p * (mu1^3 + 3*mu1*sigma1^2) + (1-input$p) * (mu2^3 + 3*mu2*sigma2^2), 
                         4)
           )
})
```

```{r}
renderValueBox({
  mu1 = sqrt((1-input$p)/input$p) * input$radius * sin(input$angle)
  mu2 = -sqrt(input$p/(1-input$p)) * input$radius * sin(input$angle)
  sigma1 = sqrt(1/input$p) * input$radius * cos(input$angle)
  sigma2 = sqrt(1/(1-input$p) * (1-input$radius^2))

  valueBox(value = round(input$p * (mu1^4 + 6*mu1^2*sigma1^2 + 3*sigma1^4) + (1-input$p) * (mu2^4 + 6*mu2^2*sigma2^2 + 3*sigma2^4) - 3, 
                         4)
           )
  })
```



